\vspace{-2cm} \LARGE Divergence-Free Smoothed Particle Hydrodynamics \\ \large Advanced Simulation and Visualization of Fluids in Computer Graphics

### 2017-01-12 report/paper.tex

    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
    The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
    The second part of Equation~\ref{eq:nsDensity}, $\nabla \cdot \textbf{v} = 0$, induce that the velocity field is free from divergence.
    From the continuity equation $\frac{D\rho}{Dt} = -\rho \nabla \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
    From this it follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
    In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
    The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
    To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
    Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.

@@

% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
In order to assure a stable simulation it is necessary to adapt the time step depending on the motion.
The quicker the particles move the smaller time steps are taken such that the numerical 
error of the explicit time integration is negligible.
For this is a condition proposed by 
Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
the maximum velocity $\vect{V_{max}}$ in the scene.

\begin{equation} \label{eq:cfl}
\Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
\end{equation}

@@

\subsection{Neighbourhood search} \label{section:neighbour}
    Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
    Searching through all particles for neighbours within the cutoff distance $h$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
    The cutoff distance $h$ is the kernels smoothing radius.
    To make this faster a cell list was implemented.
    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $h$.
    Each cell spans on a surface in space.
    A particle belongs to a cell if it is inside the space the cell occupy.
    When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see figure~\ref{fig:cellList}.

@@

    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the index of the particles that belong to the cell.
    The amount of cells are decided by dividing the scene into cells of length $h$.
    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_{min}$ is the lowest position of the scene space.
    If the particle moves out of its cell it is reassigned to the new cell it is inside.

@@

    The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
    After the symmetric pressure force has been applied to the new velocity the density change in Equation \ref{eq:dpdt} should be equal or close to zero.
    If the error is too large the process will be iterated until desired result have been achieved.

>>>>

    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
    The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
    The second part of Equation~\ref{eq:nsDensity}, $\nabla \cdot \textbf{v} = 0$, induce that the velocity field is free from divergence.
    From the continuity equation $\frac{D\rho}{Dt} = -\rho \nabla \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
    From this it follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
    In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
    The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
    To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
    Section~\ref{densitySolver} and~\ref{divergenceSolver} give a deeper description of the solvers used in the project.

@@

% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
In order to assure a stable simulation it is necessary to adapt the time step depending on the motion.
The quicker the particles move the smaller time steps are taken such that the numerical 
error of the explicit time integration is negligible.
A condition was proposed by 
Courant$-$Friedrichs$-$Lewy, often refered to as the CFL-condition, which is shown in Equation \ref{eq:cfl}.
The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
the maximum velocity $\vect{V}_{max}$ in the scene.

\begin{equation} \label{eq:cfl}
\Delta t \leq \frac{0.4 \cdot d}{\vect{V}_{max}}
\end{equation}

@@

\subsection{Neighbourhood search} \label{section:neighbour}
    Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles' neighbours.
    Searching through all particles for neighbours within the cutoff distance $h$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
    The cutoff distance $h$ is the kernels' smoothing radius.
    To make this faster a cell list was implemented.
    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $h$.
    Each cell spans on a surface in space.
    A particle belongs to a cell if it is inside the space the cell occupy.
    When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see Figure~\ref{fig:cellList}.

@@

    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the $x$, $y$ and $z$ coordinates and the fourth dimension is for storing the index of the particles that belong to the cell.

    The amount of cells are decided by dividing the scene into cells of length $h$.
    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_{min}$ is the lowest position of the scene space.
    If the particle moves out of its cell it is reassigned to the new cell it is inside.

@@

    The stiffness parameters need to be determined iteratively since neighboring particles depend on each other.
    After the symmetric pressure force has been applied to the new velocity should the density change in Equation \ref{eq:dpdt} be equal or close to zero.
    If the error is too large the process will be iterated until desired result have been achieved.
